\section{Publications and Awards}
\label{sec:publications}

The contributions of this thesis have been presented to the scientific community 
in a series of international forums, such as: journals and conferences.

\subsection{International \acs{jcr} Journals}

% The current versions of the AdaptUI mobile user interface adaptation 
% system were published in the following journals:

\begin{itemize}
  \item Castillejo, E., Almeida, A., {López-de-Ipiña}, D., 2014. Ontology Based 
  Model for Supporting Dynamic and Adaptive User Interfaces. International 
  Journal of Human-Computer Interaction 30, 771–786. 
  doi:10.1080/10447318.2014.927287.


  \item Eduardo Castillejo, Aitor Almeida, Diego {López-de-Ipiña} and Liming Chen. 
  Modeling Users, Context and Devices for Ambient Assisted Living 
  Environments. Sensors, MDPI. vol. 14, no. 3, pp. 5354-5391, \acs{doi}: 
  10.3390/s140305354, \acs{issn} 1424-8220, \acs{jcr} Impact Factor (2012): 1.953, 
  January 2014.
  
  \item Eduardo Castillejo, Aitor Almeida, and Diego {López-de-Ipiña}. Modelling 
  users, context and devices for adaptive user interface systems. International 
  Journal of Pervasive Computing and Communications 10, no. 1 (January 2014): 
  5-5
\end{itemize}
  
As mentioned before, this dissertation arises from the weaknesses and future work
identified in previous research works. 

\begin{itemize}
  \item Aitor Almeida, Pablo Orduña, Eduardo Castillejo, Diego {López-de-Ipiña} 
  and Marcos Sacristan. A method for automatic generation of fuzzy membership 
  functions for mobile device’s characteristics based on Google Trends. 
  Computers in Human Behaviour (Journal). Volume 29, Issue 2. Pages 510–517. 
  Impact Factor (2011): 2.293 \acs{doi}: 10.1016/j.chb.2012.06.005. March 2013.
  
  \item Aitor Almeida, Pablo Orduña, Eduardo Castillejo, Diego {López-de-Ipiña}, 
  Marcos Sacristán. Imhotep: an approach to user and device conscious mobile 
  applications. Personal and Ubiquitous Computing (Journal). Volume 15, Issue 
  4, pp 419–429. Springer. Impact Factor (2009): 1.554. \acs{issn}: 1617-4909. 
  \acs{doi}: 10.1007/s00779-010-0359-8. January 2011.
\end{itemize}


\subsection{International Conferences}

As this thesis started as a result of previous research work related to user 
interface adaptation, the following publications are included:

\begin{itemize}
  \item Aztiria, A., Castillejo, E., Almeida, A., {López-de-Ipiña}, D., 2014. 
  Adapting user interfaces based on user preferences and habits, in: 
  Proceedings of the 10th International Conference on Intelligent Environments 
  (IE14). Presented at the 10th international conference on Intelligent 
  Environments (IE14), Shangai, China.


  \item Aitor Almeida, Pablo Orduña, Eduardo Castillejo, Diego {López-de-Ipiña}, 
  and Marcos Sacristán. An approach to automatic generation of fuzzy 
  membership functions using popularity metrics. In Information Systems, 
  E-learning, and Knowledge Management Research, pp. 528-533. Springer Berlin 
  Heidelberg, 2013.
  
  \item Aitor Almeida, Pablo Orduña, Eduardo Castillejo, Diego {López-de-Ipiña}, 
  Marcos Sacristán. Adaptative applications for heterogeneous intelligent 
  environments. ICOST 2011: 9th International Conference on Smart Homes and 
  Health Telematics. Montréal, Canada, June 2011. LNCS6719, Toward Useful 
  Services for Elderly and People with Disabilities, Springer, \ac{isbn}: 
  978-3-642-21534-6, pp. 1-8.
  
  \item Aitor Almeida, Pablo Orduña, Eduardo Castillejo, Diego {López-de-Ipiña}, 
  Marcos Sacristán. A user-centric approach to adaptable mobile interfaces 
  Actas del II International Workshop of Ambient Assisted Living (IWAAL 2010), 
  p.p. 153-160 Valencia, Spain, September 7-10, 2010 (\ac{isbn}: 978-84-92812-67-7).
\end{itemize}


\subsection{Awards}
\label{sec:awards}

AssitedCity, the use case developed using the Imhotep framework for adaptive
user interfaces was awarded in March 7, 2012 with the Via Inteligente 
award\footnote{http://www.viainteligente.com/premios2012.html}. This prize was 
possible thank to the collaboration with Dr. Aitor Almeida and Dr. Pablo Orduña, 
and thank to the supervision of Dr. Diego López-de-Ipiña. 
% The previous work carried out with Dr. Aitor Almeida, Dr. Pablo Orduña and  
% with the Imhotep framework was awarded in March 7, 2012 with the Via Inteligente 
% award.