\subsubsection{2005: Yamabe et al.: The Citron Framework}
%\cite{yamabe_citron:_2005}
\label{sec:yamabe}

In 2005~\citet{yamabe_citron_2005} present a framework for personal devices 
which gathers context information about the user and his/her surrounding 
environment. Muffin, a personal sensor-equipped device is designed. Using it, 
several context parameters are gathered.

Sensor information is also considered for evaluating several user high-level 
context information. For example, accelerometer readings might recognize a 
walking or running activity, shaking and rotating, an so forth. The microphone 
is not only used to measure the ambient noise. It is also useful for detecting 
the place where the user is (e.g., meeting room, restaurant, street and so on).

The context acquisition is categorized in two different groups: the user and 
the environment. For the user several issues are analysed. For example, 
activity recognition requires the user to use the device in specific ways (it 
is not the same to use it with hands or waist-mounted). Another problem they 
encounter is about the time consuming process since an event is captured, then 
processed and finally validated. The last contextual issue deals with the 
intrinsic complexity and ambiguity of context information. For example, the 
meaning of what is loud might depend on the current situation (e.g., if the 
user is in a meeting room, if the user is in the street, and so on). For the 
environment, Muffin suffer several heat problems due to the sensors sensitivity
to environment temperature. This way, sometimes the gathered measures are invalid.

% \begin{table}
% \caption{Modeled context parameters \cite{yamabe_citron:_2005}}
% \label{tbl:yamabe}
% \begin{tabular}{ll}
% Sensor category & Modeled parameters  \\
% \hline
% Environmental sensors & Air temperature \\
%  & Relative humidity \\ 
%  & Barometer \\
%  
% Physiological sensors & Alcohol gas \\
%  & Pulse \\
%  & Skin temperature \\
%  & Skin resistance \\
% 
% Motion/location sensors & Compass/tilt \\
%  & 3D accelerometer \\
%  & Grip \\
%  & Ultrasonic range finder \\
%  & GPS \\
% 
% Other sensors & RFID tags \\
%  & Front rear images \\
%  & Sound \\
% \end{tabular} 
% \end{table}