\section{The AdaptUI Rules Set}
\label{sec:adaptui_rules}

As said in the introduction of this chapter, the developed adaptation platform is 
supported by two bases. The first one, the AdaptUIOnt ontology, described in the 
previous sections. The second one, a set of rules that uses the knowledge represented
by the ontology and triggers several actions aiming the user interface adaptation. 
In this dissertation the adaptation process is understood as a three step procedure: 

\begin{itemize}
 \item First, the knowledge of the main entities needs to be collected. Once it
 is gathered, a normalization process is performed. This process classifies the 
 collected information so it can be easily read. For example, a noise sensor
 might sense a 100~\acp{db} sound. Initially, the platform might not understand
 this value. Thus, using several ranges explained during this chapter (see 
 Table~\ref{tbl:luminance}, Table~\ref{tbl:sounds} and Table~\ref{tbl:batteries})
 a high level classification is carried out.
 
 \item Secondly, and taking into account the transformed knowledge in the previous
 step, several actions are performed to adapt the user interface to the current 
 situation (represented through the ontology in the previous stage).
 
 \item Finally, a refinement of the adapted user interface is available to improve
 the final results.
\end{itemize}

This conceptual division of the adaptation process in the adaptation platform 
results in a three set of different rules, which are detailed below.


\subsection{The Pre-adaptation Rules}
\label{sec:preadaptation_rules}

This group of rules are designed to check those values which come from each entity 
of the Entities Model in order to translate them to the auxiliary models. For 
example, a 35,000 \ac{lx} context brightness indicates the amount of \ac{lx} of the 
current environment. This value is then classified to a more verbose subgroup 
(in this case, \textit{direct\_sunlight}); a 10 device battery value represents 
the percentage of the remaining battery. This is translated as a 
\textit{non\_sufficient} battery level in the auxiliary device class. Thus, the
corresponding adaptation rule checks these new parameters and decides whether the
adaptation should follow one path or the other. In this dissertation we consider
that 10\% of the total battery levels for a device might not be enough to run an
adaptation process. Of course this is not accurate. The user might be able to
use a nearby power adaptor. Besides, 10\% battery level might not be enough if
the user aims to browse the web, make a video call, and so forth, but it might be
sufficient carrying out one adaptation. This means that these values are established
by the developer regarding the just the adaptation, but they directly depend on
the user context. This set of rules directly affects the \textit{UserAux},
\textit{ContextAux} and \textit{DeviceAux} classes, being the main reason for
these classes to belong to the Dynamic Model introduced in Section~\ref{sec:dynamic_model}.

Table~\ref{tbl:pre_adaptation_rules} describe the default pre-adaptation rules
included in AdaptUIOnt.

\begin{table}
  \caption{The AdaptUIOnt pre-adaptation rules.}
 \label{tbl:pre_adaptation_rules}
\footnotesize
\centering
 \begin{tabular}{l l l}
  \hline 
  \textbf{Pre-adaptation rule} 			& \textbf{Involved ontologies} 	& \textbf{Description} 	\\
						& \textbf{and classes} 		& 			\\
  \hline
  \textit{checkBatteryLevelIsSufficient}& \textit{soupa:Battery}	& Considering Table~\ref{tbl:batteries} this rule\\
					& \textit{adaptui:Device}	& evaluates if the current battery 		\\
					& \textit{adaptui:DeviceAux}	& level is enough to perform any 		\\
					&				& adaptation. 					\\
  \hline
  
  \textit{checkLightLevelDarkOvercast}	& \textit{adaptui:Context}	& These rules evaluate the context   		\\
  \textit{checkLightLevelTwilight}	& \textit{adaptui:ContextAux}	& light input through sensors by  		\\
  \textit{checkLightLevelSunrise}	& \textit{gumo:Light}		& using the classification shown in 		\\
  \textit{checkLightLevelOvercast}	& 				& Table~\ref{tbl:luminance}. The result of this \\
  \textit{checkLightLevelOfficeLightning}&				& evaluation would be a verbose  		\\
  \textit{checkLightLevelOfficeHallway}	&				& concept (e.g., \textit{sunrise}) which is   	\\
  \textit{checkLightLevelMoonlessClear}	&				& stored in the ontology using the  		\\
  \textit{checkLightLevelDayLight}	& 				& \textit{contextAuxHasLightLevel} datatype	\\
  \textit{checkLightLevelDirectSunlight}& 				&  property.					\\
  \textit{checkLightLevelFullMoon}	& 				& 						\\
  \textit{checkLightLevelLivingRoom}	& 				& 						\\
  \hline
  
  \textit{checkNoiseLevelBreathing}	& \textit{adaptui:Context}	& These rules evaluate the context 	 	\\
  \textit{checkNoiseLevelTruck}		& \textit{adaptui:ContextAux}	& noise input through sensors by  		\\
  \textit{checkNoiseLevelTrain}		& \textit{gumo:Noise}		& using the classification  shown in 		\\
  \textit{checkNoiseLevelTraffic}	& 				& Table~\ref{tbl:sounds}. The result of this 	\\
  \textit{checkNoiseLevelOffice}	&				& evaluation would be a verbose 		\\
  \textit{checkNoiseLevelLibrary}	&				& concept (e.g., \textit{sunrise}) which is  	\\
  \textit{checkNoiseLevelLeavesMurmuring}&				& stored in the ontology using the 		\\
  \textit{checkNoiseLevelBreathing}	&				& \textit{contextAuxHasNoiseLevel} datatype	\\
  \textit{checkNoiseLevelBuildingWork}	&				& property.					\\
  \textit{checkNoiseLevelConversation}	&				& 						\\
  \textit{checkNoiseLevelFactory}	&				& 						\\
  \textit{checkNoiseLevelGig}		&				& 						\\
  \textit{checkNoiseLevelJackhammer}	&				& 						\\
  \hline
  
  \textit{checkUserHasAttentionRestriction}& \textit{adaptui:Activity}	& If the current activity impedes the 		\\
  \textit{checkUserHasNoRestriction}	& \textit{adaptui:UserAux}	& user, these rules would store the 		\\
  \textit{checkUserHasSightRestriction}	&				& corresponding boolean value in the 		\\
  \textit{checkUserHasMovementRestriction}&				& \textit{userAuxHasRestriction} datatype 	\\
  \textit{checkUserHasHandsRestriction}	&				& property.					\\
  \textit{checkUserHasHearingRestriction}&				& 						\\
  \hline
\end{tabular}
\end{table}

Equation~\ref{ec:pre_adaptation_rule} shows an example of the 
\textit{checkNoiseLevelTraffic} pre-adaptation rule.

\footnotesize
\begin{equation} \label{ec:pre_adaptation_rule}
  \begin{align*} 
  Context(?context) \& Noise(?noise) \& ContextAux(?caux) \& \\
  contextHasNoise(?noise, ?value) \& lessThanOrEqual(?value, 70) \& \\
  greaterThan(?value, 60) \\
  \Rightarrow \\
  contextAuxHasNoiseLevel(?caux, "traffic")
  \end{align*}
\end{equation}
\normalsize


\subsection{The Adaptation Rules}
Depending on the checked rules and the auxiliary classes’ status, different 
rules are triggered. These rules result in different values for the \textit{Adaptation} 
class, which is the class queried in the device platform for bringing the final 
adaptation to the device. Table~\ref{tbl:adaptation_rules} details the default
adaptation rules in the AdaptUIOnt ontology.

\begin{table}
  \caption{The AdaptUIOnt adaptation rules.}
 \label{tbl:adaptation_rules}
\footnotesize
\centering
 \begin{tabular}{l l l}
  \hline 
  \textbf{Adaptation rule} 	& \textbf{Involved ontologies} 	& \textbf{Description} 		\\
				& \textbf{and classes} 		& 				\\
  \hline
  \textit{adaptBrightness}	& \textit{adaptui:Adaptation}	& These rules consider the  	\\
  and				& \textit{adaptui:DeviceAux}	& values stored in the 		\\
  \textit{adaptVolume}		& \textit{adaptui:UserAux}	& \textit{UserAux} class to adapt the \\
				& \textit{adaptui:Brightness}	& brightness and volume 	\\
				& \textit{adaptui:Volume}	& accordingly. 			\\
				& \textit{adaptui:Display}	& 				\\
  \hline 
  \textit{adaptButtonSize}	& \textit{adaptui:Adaptation}	& These rules take into account	\\
				& \textit{adaptui:DeviceAux}	& the current button configu-	\\
				& \textit{adaptui:UserAux}	& ration and the sensed context	\\
  \textit{adaptButtonColor}	& \textit{adaptui:Button}	& disabilities.			\\
  \textit{adaptButtonBackgroundColor}&				& 				\\
  \textit{adaptButtonTextColor}	& 				& 				\\

  \hline
  \textit{adaptEditTextSize}	& \textit{adaptui:Adaptation}	& These rules take into account	\\
				& \textit{adaptui:DeviceAux}	& the current edit text configu-\\
				& \textit{adaptui:UserAux}	& ration and the sensed	context	\\
  \textit{adaptEditTextBackgroundColor}	& \textit{adaptui:EditText}& disabilities.		\\
  \textit{adaptEditTextTextSize}& 				& 				\\
  \textit{adaptEditTextTextColor}&				& 				\\

  \hline 
  \textit{adaptTextViewtSize}	& \textit{adaptui:Adaptation}	& These rules take into account	\\
				& \textit{adaptui:DeviceAux}	& the current text view configu-\\
				& \textit{adaptui:UserAux}	& ration and the sensed	context	\\
  \textit{adaptTextViewtBackgroundColor}& \textit{adaptui:TextView}& disabilities.		\\
  \textit{adaptTextViewtTextSize}&				& 				\\
  \textit{adaptTextViewtTextColor}&				& 				\\
  \hline 
\end{tabular}
\end{table}

Equation~\ref{ec:adaptation_rule} shows an example of the \textit{adaptVolume} 
adaptation rule.

\footnotesize
\begin{equation} \label{ec:adaptation_rule}
  \begin{align*} 
  Adaptation(?adaptation) \& DeviceAux(?device) \& ContextAux(?context) \& \\
  deviceAuxBatteryIsSufficient(?device, ?battery) \& equal(?battery, true) \& \\
  contextAuxHasNoise(?context, ?noise) \& equal(?noise, ``traffic'')\\ 
  \Rightarrow \\
  adaptationVolumeHasValue(?adaptation, 7)\\
  \end{align*}
\end{equation}
\normalsize


\subsection{The Usability Rules}
In order to check the usability satisfaction of the user with the provided
adapted user interface, a usability rules set is provided. By checking several
usability metrics (detailed in Section~\ref{sec:usability_metrics}) these rules
determine if the interaction with the adapted user interface might be considered
enough. Table~\ref{tbl:usability_rules} introduced the usability rules included
in AdaptUIOnt.

\begin{table}
  \caption{The AdaptUIOnt usability rules. The metrics mentioned in this table
  are detailed in Table~\ref{tbl:effectiveness_metrics} and Table~\ref{tbl:productivity_metrics}.}
 \label{tbl:usability_rules}
\footnotesize
\centering
 \begin{tabular}{l l l}
  \hline 
  \textbf{Usability rule} 	& \textbf{Involved ontologies} 	& \textbf{Description} 		\\
				& \textbf{and classes} 		& 				\\
  \hline
  \textit{checkTaskEffectiveness}&\textit{adaptui:UserAux}	& It measures the proportion of \\
				& \textit{adaptui:Effectiveness}& goals of the task achieved	\\
				& \textit{adaptui:Polisher}	& correctly. 			\\
  \textit{checkTaskCompletion}	& 				& It measures the proportion of	\\
				& 				& the task that is completed.	\\
  \textit{checkErrorFrequency}	& 				& It measures the frequency of 	\\
				& 				& errors.			\\
  \hline
  \textit{checkTaskTime}	& \textit{adaptui:UserAux}	& It measures the required time	\\
				& \textit{adaptui:Productivity}	& to complete the current task.	\\
  \textit{checkTaskEfficiency}	& \textit{adaptui:Polisher}	& It measures the efficiency of	\\
				& 				& the user.			\\
  \textit{checkEconomicProductivity} &				& It measures the cost-effectiveness\\
				&				& of the user.			\\
  \textit{checkProductiveProportion} & 				& It measures the proportion of \\
				& 				& the time the user is performing\\
				&				& productive actions.		\\
  \textit{checkRelativeUserEfficiency}& 			& It compares the efficiency of the\\
				&				& user compared to an expert.	\\
  \hline
\end{tabular}
\end{table}

Equation~\ref{ec:usability_rule} shows an example of the \textit{checkRelativeEfficiency} 
adaptation rule.

\footnotesize
\begin{equation} \label{ec:usability_rule}
  \begin{align*} 
  UserAux(?user) \& Productivity(?productivity) \& Polisher(?polisher) \&\\ 
  userAuxHasProductivityMetrics(?user, ?productivity) \& \\
  hasRelativeEfficiency(?productivity, ?efficiency) \& \\
  lessThanOrEqual(?efficiency, 0.5) \& \\
  \Rightarrow \\
  launchPolisherRules(?polisher, true)
  \end{align*}
\end{equation}
\normalsize

\subsection{The Post-adaptation Rules}
Once the user interface is adapted, a concrete architecture module (see 
Section~\ref{sec:adaptation_polisher}) monitors the user activity. Hence, through
a series of usability metrics the adaptation is considered satisfactory by
the adaptation platform. If it detects that the usability level is inadequate,
these rules are triggered, changing the user model.

\begin{table}
  \caption{The AdaptUIOnt post-adaptation rules. For edit texts and text views
  the same rules that are applied for the buttons are provided.}
 \label{tbl:post_adaptation_rules}
\footnotesize
\centering
 \begin{tabular}{l l l}
  \hline 
  \textbf{Adaptation rule} 	& \textbf{Involved ontologies} 	& \textbf{Description} 		\\
				& \textbf{and classes} 		& 				\\
  \hline
  \textit{incrementBrightness}	& \textit{adaptui:UserAux}	& It increments the brightness of\\
				& \textit{adaptui:Polisher}	& the device in 1F.		\\
  \textit{decrementBrightness}	& \textit{adaptui:Adaptation}	& It decrements the brightness of\\
				& 				& the device in 1F.		\\
  \textit{incrementVolume}	& 				& It increments the volume of	\\
				& 				& the device in 1 unit.	\\
  \textit{decrementVolume}	& 				& It decrements the volume of	\\
				& 				& the device in 1 unit.	\\
  \hline
  \textit{incrementButtonSize}	& \textit{adaptui:UserAux}	& It increments the size of the \\
				& \textit{adaptui:Polisher}	& button adding 10 dpis. 	\\
  \textit{decrementButtonSize}	& \textit{adaptui:Adaptation}	& It decrements the size of the \\
				& 				& button in 10 dpis.		\\
  \textit{darkenButtonBackgroundColor}&				& It darkens the colour.	\\
  \textit{lightenButtonBackgroundColor}&			& It lightens the colour.	\\
  \textit{darkenButtonTextColor}&				& It darkens the text colour.	\\
  \textit{lightenButtonTextColor}&				& It lightens the text colour.	\\
  \hline
\end{tabular}
\end{table}

Equation~\ref{ec:post_adaptation_rule} shows an example of the \textit{incrementButtonSize} 
post-adaptation rule.

\footnotesize
\begin{equation} \label{ec:post_adaptation_rule}
  \begin{align*} 
  Polisher(?polisher) \& launchPolisherRules(?polisher, true) \&\\
  UserAux(?user) \& userAuxHasEffectivenessMetrics(?user, ?effectiveness) \& \\
  effectivenessMetricHasErrorFrequency(?effectiveness, ?freq?) \&\\
  greaterThan(?freq, 0.5) \& Adaptation(?adaptation) \&\\
  adaptationHasButtonSize(?adaptation, ?size)
  \Rightarrow \\
  adaptationButtonHasSize(?adaptation, ?size + 10)
  \end{align*}
\end{equation}
\normalsize



The given set of rules by the adaptation platform have been included as they are
understood as vital for a precise adaptation. Nevertheless, and as it will be
explained in Chapter~\ref{cha:architecture}, these rules are not final. This
means that they are modifiable. To this end, a series of tools for developers
that allow them to change the knowledge managed by the platform are provided.


\subsection{Conclusions}
\label{sec:rules_conclusions}

In this second part of the chapter the second main basis of the adaptation platform
has been presented: the set of rules. Divided into three different groups
(considering that the adaptation process is understood as a three step procedure)
these rules aim to finally generate a user interface adaptation under the following
steps:

\begin{itemize}
  \item Firstly, a classification and normalization of the collected knowledge of
  the main entities is needed. This might be low level knowledge. Thus, a higher
  level information is obtained during this process. To reach this goal, the
  pre-adaptation rules set is needed. These rules affect the \textit{UserAux},
  \textit{ContextAux} and \textit{DeviceAux} classes.
  
  \item Next, once the knowledge has been normalized, the adaptation rules set
  are triggered, generating the corresponding changes in the \textit{Adaptation}
  class in the AdaptUIOnt ontology.
  
  \item Finally, as the adaptation might be improvable, several rules are executed
  regarding a refinement of the last adapted user interface. This is detailed
  in Section~\ref{sec:adaptation_polisher}.
\end{itemize}