%: ----------------------- introduction file header -----------------------
\begin{savequote}[50mm]
So you see, the only proof of what you are is in the way you hear the truth.
% Don't be scared, live to win, although they're always gonna tell you it's a sin.
\qauthor{Motörhead - Stay Clean}
\end{savequote}

% the code below specifies where the figures are stored
\ifpdf
\graphicspath{{5_experiments_and_results/figures/PNG/}
{5_experiments_and_results/figures/PDF/}{5_experiments_and_results/figures/}}
\fi

\chapter{Evaluation}
\label{cha:evaluation}
%------------------------------------------------------------------------- 

% INTRODUCTION OF THE CHAPTER

In the previous chapters several contributions have been presented. All these
contributions together define AdaptUI, a mobile platform which regards adaptive 
and dynamic user interfacing for mobile devices. AdaptUI combines mobile 
semantics and reasoning, supported by an ontology which models user's 
characteristics, context current conditions and device's hardware and software 
characteristics. In this chapter an evaluation of the designed ontology and the 
semantic based platform which manages it is performed. This evaluation is 
addressed through several experiments, regarding cycle time responsiveness, 
system performance with specific operations (e.g., reasoning over a controlled 
set of rules), other alternatives and, finally, analysing users' acceptance and 
feedback.

Consequently, this chapter is divided in two different parts considering the 
experiments' goals. On the one hand, an evaluation of the technical performance
of the AdaptUI system is presented. Including quantitative and qualitative 
experiments, this technical evaluation has been developed as follows:

\begin{enumerate}[label=\alph*)]
  \item First, the developed mobile adaptation of the Pellet reasoning engine 
  is tested through several scenarios. In these scenarios different   
  modifications of the AdaptUIOnt ontology are presented in order to evaluate 
  the performance differences of \textit{Pellet4Android} and Pellet.
  
  \item Next, as the system goal is to adapt the user interface, a comparison 
  with another user interface adaptation solution has been performed: the 
  Imhotep framework. Imhotep is a user interface adaptation framework which 
  aims to ease the development of adaptable and more accessible user 
  interfaces.
  
  \item Finally, several scenarios are presented in which the adaptation process
  of the AdaptUI system is described step by step, from the scenario conditions
  to the adaptation results. These scenarios have been divided into three 
  different groups, each group aiming to cover different types of adaptation 
  and environments. This classification includes:
  \begin{itemize}
    \item A scenario in which the limitations suffered by the user are caused
    by the current context conditions. 
    
    \item A second scenario introducing several limitations due to the  
    activities the user performs.
    
    \item A final case in which several disabilities impedes the user to 
    interact properly with the user interface.
  \end{itemize}
\end{enumerate}

On the other hand, an evaluation regarding users, without technical knowledge
requisites, is performed. This part of the evaluation is based on the user
satisfaction when using AdaptUI. The scenario for this experiment includes the
user and an Android based tutorial through which the user indicates several 
interaction parameters. Taking into account different context situations, 
several applications are adapted consequently. After the adaptation, the user 
is asked to complete a questionnaire of usability. More concretely, the \ac{sus} 
questionnaire is presented (among several extra questions to classify the 
results).

Hence, the the reminder of this chapter has been structured as follows: 

\begin{itemize}
  \item A technical evaluation, which experiments are detailed in 
  Section~\ref{sec:technical_evaluation}. These experiments regard the 
  reasoning performance comparing Pellet for desktop versus Pellet for mobile 
  devices. The provided \acp{api} have been also evaluated in an experiment
  including developers.
  
  \item A usability evaluation of the system, which details are accordingly 
  described in Section~\ref{sec:user_evaluation}. This evaluation includes 
  users interacting with the AdaptUI system. In order to evaluate the usability 
  of AdaptUI a usability questionnaire is presented to the users. 
  Section~\ref{sec:sus_results} presents Table~\ref{tbl:additional_questions} and
  Table~\ref{tbl:sus_questionnaire_results}, which shows the responses from all
  the users participating in the evaluation of AdaptUI.

  \item Finally, Section~\ref{sec:evaluation_discussion} and 
  Section~\ref{sec:evaluation_conclusions} summarize and analyse the results
  obtained and detailed in the previous sections, regarding both technical and
  usability aspects.
\end{itemize}

\input{5_experiments_and_results/51_technical_evaluation}
\input{5_experiments_and_results/512_imhotep_comparison}
\input{5_experiments_and_results/513_scenarios}
\input{5_experiments_and_results/514_developers}
%
\input{5_experiments_and_results/52_user_evaluation}
% 
\input{5_experiments_and_results/53_evaluation_discussion}
\input{5_experiments_and_results/54_evaluation_conclusions}