\begin{savequote}[50mm]
You take the blue pill, the story ends, you wake up in your bed and believe 
whatever you want to believe. You take the red pill, you stay in Wonderland, 
and I show you how deep the rabbit hole goes.
% We are marching to a land far from home. No one can say who will return.
% I feel the hard fight has begun, no more living on the run.
% I feel a hard fight has begun, I know it's been done.
% Yes, a hard fight has begun.
\qauthor{The Matrix}
\end{savequote}

\chapter{Introduction}
\label{cha:introduction}

% the code below specifies where the figures are stored
\ifpdf
    \graphicspath{{1_introduction/figures/PNG/}{1_introduction/figures/PDF/}{1_introduction/figures/}}
\else
    \graphicspath{{1_introduction/figures/EPS/}{1_introduction/figures/}}
\fi

%------------------------------------------------------------------------- 

Over the years we are witnessing the growth of new and heterogeneous mobile and
smart devices and services with a wider range of possibilities. Faster processors,
larger memories and more accessible sensor capabilities have allowed the community to
develop context-aware applications which, taking into account the user preferences
and the context situation, can be customized for the final user. This usually
results into the same application having different behaviours, aspects and 
available features depending on the specific target user. Besides, the spread 
of intelligent environments provides relevant information about the current 
context of the agents and involved entities. As a consequence, local 
governments and public administrations have discovered the importance of 
working with context data~\citep{caragliu_smart_2009}. Thus, they try to 
improve cities infrastructures and citizens satisfaction.

This situation has brought the appearance of new research domains. This 
dissertation focuses on one of these domains: adaptive user interfaces. 
Adaptive user interfaces arise from the need to cover a wider range of users 
and environment conditions. This area is related to different research domains 
or trends. For instance, universal or inclusive design. Universal design refers 
to a set of guidelines for producing different kind of environments (i.e., 
buildings, software applications and any kind of product) accessible and usable 
to both people with and without disabilities and dependant people (as the 
elderly).

During this dissertation we have studied how each user has his own preferences,
even those who suffer from similar disabilities. We have also found that usual 
market applications and devices are usually unable to guarantee a comfortable 
interaction in many situations for these users. Although several accessibility 
tools are provided (related to smart devices), these users prefer non smart 
devices due to the physical interaction that they provide (as physical touch 
buttons, not touchable screens). This kind of devices are easy to use and the 
feedback that they provide is also easy to understand. Hence, this situation has
revealed a lack of substantial efforts in the adaptive user interfaces domain. 
This makes this group of users suffer from interaction inattention.

Hardware advances have brought haptic displays, high definition, curved 
screens and cameras, accelerometers and different kind of sensors, faster 
connectivity capabilities, and so forth. On the other hand, software development
allow us to use Internet services as we did with a computer. Nevertheless,
this progress does not reflect the current society requirements (including
inclusive design). While these advances keep reaching the ubiquitous future
there are several groups of people that suffer from inattention: the elderly and
people with disabilities. These users have special needs. The elderly usually
have mobility, sight and hearing problems, while the disabled suffer from more
specific and usually severe impairments. There are several accessibility tools
that try to reduce the existing interaction boundaries between disabled users
and devices~\citep{gregor_designing_2002}\citep{burgstahler_designing_2002}.
For example, the Android operating system allows developers to build more accessible 
applications by using custom controls and interaction alternatives (e.g., 
gestures)\footnote{http://developer.android.com/guide/topics/ui/accessibility/index.html}.
Apple's iOS uses zoom, larger text, colours inversions, dictation and voice
control to avoid interaction issues\footnote{https://www.apple.com/accessibility/ios/}.
Regrettably, these tools are too static. For example, they do not take the user
context into account and they do not understand and learn from the users'
experiences. Besides, they are more adaptable than adaptive. Adaptive systems
have the ability to dynamically adapt themselves to the current task and user.
On the other hand, adaptable systems' functionalities are changed with user
intervention~\citep{fischer_user_2001}.

The fact is that user interface adaptation has evolved in the latests 20 years.
From the simplest preferences (e.g., modifying the resolution of a screen or
monitor) to more sophisticated solutions (e.g., smartphones' brightness automatic
controls), the community has attempted to customize applications as far as 
possible. These adaptations have grown in complexity, covering a wider range of 
users, as well as taking into account the current context situation. Therefore, 
context and user modelling have become a real challenge in this domain. This is 
because of the context variability and the set of different capabilities that 
users can have.

However, current mobile devices offer new possibilities due to their 
computational capabilities. Hence, during the following chapters we present a
series of contributions which are grouped together forming AdaptUI, a user 
interface adaptation framework which has a twofold purpose: to reduce the 
usability problems when users interact with their devices, and to encourage 
developers to include adaptive engines in their applications to make them more 
inclusive.

\input{1_introduction/1_background}
\input{1_introduction/2_motivation}
\input{1_introduction/3_hypothesis_and_goals}
\input{1_introduction/4_context}
\input{1_introduction/5_methodology}
\input{1_introduction/6_outline}