
\section{Outline of this thesis}
\label{sec:outline}

This thesis is presented in 6 chapters (including the current one) and 3
appendices. As a guide to the organization of the remainder of this thesis:

\hspace*{5mm} \textbf{Chapter 1} outlines the motivation, hypothesis, goals and 
limitations of this research. It also describes the followed research 
methodology. \\

\hspace*{5mm} \textbf{Chapter 2} analyses the state of the art regarding user, 
context and device modelling, as well as the different adaptation solutions in 
the last 20 years. \\

\hspace*{5mm} \textbf{Chapter 3} presents AdaptUIOnt: an ontology for 
dynamic user interface adaptation specially designed for this dissertation, 
but also aiming to be used by the scientific community in the user interface 
adaptation domain.\\

\hspace*{5mm} \textbf{Chapter 4} presents the three layered architecture of 
the AdaptUI system, decoupling its modules, their purposes, and main 
characteristics.\\

\hspace*{5mm} \textbf{Chapter 5} outlines and analyses the corresponding
experiments, specially performed to evaluate the whole AdaptUI platform. This
evaluation is divided in two different parts. The first one, regarding only
technical issues of the developed system. The second one, pursuing the 
satisfaction of users.\\

\hspace*{5mm} \textbf{Chapter 6} draws the conclusions of this research work. We 
discuss the achieved goals and analyse the advantages and drawbacks of the 
developed system. Some future lines of research are proposed and the chapter 
ends with some final remarks.\\

\hspace*{5mm} \textbf{Appendix A} details the classification of the disabilities
considered for this dissertation. To do this the \ac{icf} document and the
considerations by \citet{persad_characterising_2007} are detailed.\\

\hspace*{5mm} \textbf{Appendix B} presents the default included rules in the
AdaptUIOnt, divided into the corresponding sets of rules depending on their concrete
goal.\\

\hspace*{5mm} \textbf{Appendix C} shows the results of the user evaluation
described in Chapter~\ref{cha:evaluation}.\\
