
% this file is called up by thesis.tex
% content in this file will be fed into the main document

%: ----------------------- introduction file header -----------------------
\begin{savequote}[50mm]
The lover of life's not a sinner, the ending is just a beginner.
The closer you get to the meaning, the sooner you'll know that you're dreaming.
\qauthor{Black Sabbath}
%“And upon the top of the pillars was lily work: so was the work of the pillars finished.”
%
% Bible quotes
\end{savequote}


\chapter{State of the Art}
\label{cha:state_of_the_art}

% the code below specifies where the figures are stored
\ifpdf
    \graphicspath{{2_state_of_the_art/figures/PNG/}{2_state_of_the_art/figures/PDF/}{2_state_of_the_art/figures/}}
\else
    \graphicspath{{2_state_of_the_art/figures/EPS/}{2_state_of_the_art/figures/}}
\fi

%------------------------------------------------------------------------- 

This chapter introduces several significant models and techniques related to 
adaptive user interfaces and systems during the past 20 years. Besides, user-oriented,
context-aware and device models and solutions are reviewed. Once the general
domain in which this dissertation lies has been analysed during this chapter,
the specific approaches that are close to this proposal are examined. Thus, 
the way previous proposals have addressed this dissertation's aspects will be 
analysed, highlighting their goodness and weaknesses.  

First, in Section~\ref{sec:adaptation_models} several important adaptation models
are introduced. The performed methodology and modelled parameters are analysed in
detail. Next, Section~\ref{sec:user} focuses its attention in the models
that have been used during the past 20 years for characterizing the user and his/her
capabilities. Section~\ref{sec:context} analyses the most significant
context-aware systems and the way these approaches model the context and its
characteristics. Devices are also one of the main entities that are taken into
account in this dissertation for the adaptation process. Therefore, Section~\ref{sec:devices}
examines several modelling techniques and the devices' capabilities in different
domains. Besides, a chronological review of the evolution of the presented user 
and context models will be depicted in the corresponding section.

\input{2_state_of_the_art/1_adaptation_models}
\input{2_state_of_the_art/2_user_models}
\input{2_state_of_the_art/3_context_models}
\input{2_state_of_the_art/4_device_models}
% ----------------------------------------------------------------------

