
% this file is called up by thesis.tex
% content in this file will be fed into the main document

%: ----------------------- introduction file header -----------------------
\begin{savequote}[50mm]
Personally, I think it does help, that it makes a beneficial difference, but the scientific literature on the subject is very messy.
\qauthor{Jeanne Petrek}
%“And upon the top of the pillars was lily work: so was the work of the pillars finished.”
%
% Bible quotes
\end{savequote}


\section{User Modelling}
\label{sec:user}

% the code below specifies where the figures are stored
\ifpdf
    \graphicspath{{2_state_of_the_art/figures/PNG/}{2_state_of_the_art/figures/PDF/}{2_state_of_the_art/figures/}}
\else
    \graphicspath{{2_state_of_the_art/figures/EPS/}{2_state_of_the_art/figures/}}
\fi


%------------------------------------------------------------------------- 

\subsection{A Chronological Review of the Evolution of User Models}
\label{sec:chronological_review}

Figure~\ref{fig:user_models} illustrates the chronological evolution and different 
solutions for the last 15 years. During these years different user characteristics 
have been taken into account considering the final purpose of the designed system. 

\vspace{1cm}
\setlength\taskwidth{1.7cm}

\begin{timeline}
  \label{chr:users}
  \Task[1991]{\citet{orwant_doppelgangeruser_1991}}
  \Task[1994]{\citet{brajnik1994shell}, \citet{paiva1994tagus}, \citet{kay1994toolkit}}
  \Task[1999]{\citet{pohl_logic_based_1999}}
  \Task[2001]{\citet{fischer_user_2001}, \citet{kobsa_generic_2001}.}
  \Task[2002]{\citet{gregor_designing_2002}}
  \Task[2003]{\citet{gauch_ontology_based_2003}, \citet{razmerita_ontology_based_2003}}
  \Task[2005]{\citet{hatala_ontology_based_2005}, \citet{pereira_triple_2005}, \citet{heckmann_gumogeneral_2005}}
  \Task[2007]{\citet{persad_characterising_2007}, \citet{golemati_creating_2007}}
  \Task[2008]{\citet{casas_user_2008}}
  \Task[2012]{\citet{evers_achieving_2012}, \citet{skillen2012ontological}}
\end{timeline}
\captionof{figure}{The chronological view of the evolution of remarkable user 
models considered in this dissertation.\label{fig:user_models}}


First user modelling systems started in the late eighties. 
\citet{allen_plan_based_1979},~\citet{cohen_elements_1979}, 
~\citet{perrault_speech_1978}, and \citet{rich_building_1979}\citep{rich_user_1979} 
are examples of researchers whose works inspired next user modelling approaches. 
From there, several authors started collecting different types of information 
about the users and exhibiting, for example, different kinds of adaptations to 
them~\citep{kobsa_generic_2001}.

Before overview the evolution of the user models, a definition of user modelling 
is needed. There are at least two different perspectives which might answer this 
question. One of these is related to \ac{ai}, and it considers user modelling 
as the process through which systems gather information and knowledge about 
users and their individual characteristics. Therefore, a user model is considered 
a source of information about the user of the system which contains several
assumptions about several relevant behaviour or adaptation data. However, in
this thesis the \ac{hci} research perspective is taken, which is defined by
\citet{pohl_logic_based_1999} as follows:

\begin{description}
  \item[\Defi{User Model, by \citet{pohl_logic_based_1999}}] \hfill \\
  \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  \textit{``It refers to an a-priori model of the users of a computer system that the system
  designer has in mind, or to the assumed models that users will probably develop
  of the system and the tasks they can perform using the system''}.
  \end{mdframed}
\end{description}

Nevertheless, this definition and the \ac{ai} perspective coincide in the idea that 
every system must use information about the user to be able to see and react to 
their different problems and needs and improve the system purpose. This section 
analyses the most significant user models in the past 15 years.

\subsection{User models}
\label{sec:user_models}

%sota user models
\input{2_state_of_the_art/user/2_1_orwant}
\input{2_state_of_the_art/user/2_2_fischer}
\input{2_state_of_the_art/user/2_3_gregor}
\input{2_state_of_the_art/user/2_3_II_gauch}
\input{2_state_of_the_art/user/2_4_razmerita}
\input{2_state_of_the_art/user/2_4_hatala}
\input{2_state_of_the_art/user/2_5_pereira}
\input{2_state_of_the_art/user/2_7_heckmann}
\input{2_state_of_the_art/user/2_6_persad}
\input{2_state_of_the_art/user/2_8_golemati}
\input{2_state_of_the_art/user/2_8_casas}
\input{2_state_of_the_art/user/2_9_evers}
\input{2_state_of_the_art/user/2_12_skillen}
\input{2_state_of_the_art/user/2_x_generic_systems}
\input{2_state_of_the_art/user/comparison}


