\section{Context Modelling}
\label{sec:context}
%------------------------------------------------------------------------- 
\subsection{What is Context?}
\label{sec:context_definition}

% Context is mostly defined by the definition by Dey as follows:
Context is often defined according to \citeauthor{dey_understanding_2001}
issued definition:

\begin{description}
  \item[\Defi{Context (I), by~\citet{dey_understanding_2001}}] \hfill \\
  \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  \textit{`Context is any information that can be used to characterize the situation
  of an entity. An entity is a person, place, or object that is considered 
  relevant to the interaction between a user and an application, including the 
  user and applications themselves``}.
  \end{mdframed}
\end{description}

In the past decades there were many definitions of context~\citep{adomavicius_context_aware_2011}.
Nevertheless, the one stated by~\citet{dey_understanding_2001} is one of the most
popular and extended definitions. \citeauthor{dey_understanding_2001}'s 
definition enables developers to easily enumerate those elements which take part 
in the context for a certain application domain. \citeauthor{dey_understanding_2001} stated that:

\begin{description}
  \item[\Defi{Context (II), by~\citet{dey_understanding_2001}}] \hfill \\
  \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  \textit{``If a piece of information can be used to characterize the situation 
  of a participant in an interaction, then that information is context''}.
  \end{mdframed}
\end{description}

A proposed example explains this definition: \textit{``Take the canonical 
context-aware application, an indoor mobile tour guide. The obvious entities in 
this example are the user, the application and the tour sites. We will look at 
two pieces of information – weather and the presence of other people – and use 
the definition to determine whether either one is context. The weather does not 
affect the application because it is being used indoors. Therefore, it is not 
context. The presence of other people, however, can be used to characterize the 
user's situation. If a user is travelling with other people, then the sites they 
visit may be of particular interest to her. Hence the presence of other people 
is context because it can be used to characterize the user's situation.''~\citep{dey_understanding_2001}}

By this example, it is seen how different pieces of information are analysed to
determine if they belong to what \citeauthor{dey_understanding_2001} states 
context is. These definitions are based on research experience. 
Section~\ref{sec:context_models} shows how modelling and defining context has 
evolved in the past 20 years. 

% Context management allows us to identify the conditions of the environment. This
% way, developers are able to adapt services or applications for the user taking
% into account these conditions. To do this first there is the need of gathering
% context information. Next this information has to be somehow processed and,
% finally, it will be used to personalize and contextualize the current situation.

% ----------------------------------------------------------------------

\subsection{A Chronological Review of the Evolution of Context Management}
\label{sec:chronological_review}
% \dots
Figure~\ref{fig:context_models} shows the evolution for context modelling by 
chronological order for the last 15 years. 

\vspace{1cm}
\setlength\taskwidth{1.9cm}

\begin{timeline}
  \label{chr:context}
    \Task[2000]{\citet{chen_survey_2000}}
    \Task[2001]{\citet{jameson_modelling_2001}}
    \Task[2002]{\citet{henricksen_modeling_2002}, \citet{held_modeling_2002}}
    \Task[2004]{\citet{gu_toward_2004}}
    \Task[2005]{\citet{chen_using_2005}, \citet{yamabe_citron_2005}}
    \Task[2008]{\citet{wood_context_aware_2008}}
    \Task[2011]{\citet{baltrunas_incarmusic_2011}}
    \Task[2012]{\citet{mcavoy_ontology_based_2012}}
    \Task[2013]{\citet{almeida_assessing_2012}}
\end{timeline}
\captionof{figure}{The chronological view of the evolution of remarkable context
models.\label{fig:context_models}}


\subsection{Context Models}
\label{sec:context_models}

The first definition of \textit{context-aware systems} is given by~\citet{schilit_disseminating_1994}.
\citet{dey_understanding_2001} also defines \textit{context-aware systems} as
those systems which, using context data, provide significant information and/or
services to the user where the relevancy of the given information depends on the
user task.~\citet{schmidt_there_1999} consider several issues about context
modelling. Authors emphasize the excess of abstraction about context-aware systems
and environments which causes a lack of models to be compared . Therefore they
present a working model for context-aware systems categorized into human and
physical environment factors. In 2001~\citet{jameson_modelling_2001} studies how
context-aware computing represents a challenging frontier for researchers. In
this work, information about the environment, the user current state, longer
term user properties and the user behaviour are compared in order to take the
correct adaptation decision. Several works focused on user interface adaptation
area base their processes on context changes as triggers. However, they lack of
a common model of context in their platforms~\citep{calvary_plasticity_2002}
~\citep{nilsson_model_based_2006}.
% This fact emphasizes the argument established by Schmidt et al. \cite{schmidt_there_1999}.

In the following subsections a review of the most popular context-aware systems
is presented, mainly focusing on the modelled context parameters, techniques, 
domains and dependencies. Besides, several authors define context and 
context-awareness through their own experience. All these models and definitions 
have been considered for the context model proposed in 
Chapter~\ref{cha:ontology_model}. In Section~\ref{sec:context_model_comparison} 
Table~\ref{tbl:context_comparison} shows several significant features of each 
approach, and a comparing analysis is performed.


\input{2_state_of_the_art/context/2_1_chen}
\input{2_state_of_the_art/context/2_2_jameson}
\input{2_state_of_the_art/context/2_3_henricksen}
\input{2_state_of_the_art/context/2_4_held}
\input{2_state_of_the_art/context/2_5_gu}
\input{2_state_of_the_art/context/2_6_chen}
\input{2_state_of_the_art/context/2_7_yamabe}
\input{2_state_of_the_art/context/2_8_wood}
\input{2_state_of_the_art/context/2_9_baltrunas}
\input{2_state_of_the_art/context/2_10_mcvoy}
\input{2_state_of_the_art/context/2_11_almeida}

\input{2_state_of_the_art/context/comparison}
\input{2_state_of_the_art/context/modeling_techniques}